\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[french,english]{babel}
\usepackage{cite}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, fit, backgrounds}

% ── TikZ styles ──────────────────────────────────────────────────────────────
\tikzset{
  block/.style={rectangle, rounded corners=3pt, draw=black!70, fill=blue!8,
                text width=2.5cm, align=center, font=\footnotesize,
                minimum height=0.72cm, inner sep=4pt},
  blockred/.style={block, fill=red!8},
  blockgreen/.style={block, fill=green!9},
  blockorange/.style={block, fill=orange!10},
  blockgray/.style={block, fill=gray!10},
  decision/.style={diamond, draw=black!70, fill=yellow!15, aspect=2.4,
                   font=\scriptsize, align=center, inner sep=2pt},
  arrow/.style={-{Stealth[length=5pt]}, thick, black!70},
  groupbox/.style={draw=black!30, dashed, rounded corners=5pt,
                   fill=black!3, inner sep=6pt},
}

\begin{document}

\title{AI-Assisted Radiology Reporting Through Multi-Organ Semantic Segmentation
and Automated French Report Generation: A Clinical System Deployed on Real Data
from Sub-Saharan Africa}

\author{
  \IEEEauthorblockN{Alaaeddine Bouchamla}
  \IEEEauthorblockA{ENISO, Universit\'{e} de Sousse\\Sousse, Tunisie}
  \and
  \IEEEauthorblockN{Mehrez Abdellaoui}
  \IEEEauthorblockA{ENISO, Universit\'{e} de Sousse\\Sousse, Tunisie}
  \and
  \IEEEauthorblockN{Dahas Jalel \and Luis Felipe}
  \IEEEauthorblockA{Centre de Radiologie \'{E}milie (CRE)\\Libreville, Gabon}
}

\maketitle

% ── Abstract ─────────────────────────────────────────────────────────────────
\begin{abstract}
We present an end-to-end AI-assisted radiology reporting system combining
multi-organ semantic segmentation with automated French-language report
generation, deployed as a clinical web application for the Centre de Radiologie
\'{E}milie (CRE) in Libreville, Gabon. The system ingests DICOM CT and MRI
volumes, performs modality-aware segmentation using pretrained MONAI bundles
--- 104-organ CT segmentation via TotalSegmentator \cite{wasserthal2023} and
133-structure brain MRI segmentation via UNesT \cite{yu2023} --- and applies a
CRE-specific fine-tuning layer trained on pseudo-labels automatically generated
by MedSAM \cite{ma2024} from real patient data, requiring no manual annotation.
Segmentation outputs feed directly into per-organ volumetric metrics (volume in
mL, mean Hounsfield density, anomaly flags against published normal ranges), which
ground the generation of draft \textit{comptes rendus} in French via a
retrieval-augmented generation (RAG) pipeline built from real institutional
reports. A radiologist reviews, edits, and validates each AI draft before PDF
generation, ensuring clinical safety. Validated reports are indexed back into the
knowledge base, implementing continual learning. Evaluated on the CRE pseudo-label
validation set, Stage~2 fine-tuning achieves IoU~=~[FILL], Dice~=~[FILL]. Report
quality was assessed by the clinical partner on [FILL] cases. To our knowledge,
this is the first French-language AI radiology reporting system built on real
clinical data from sub-Saharan Africa.
\end{abstract}

\begin{IEEEkeywords}
semantic segmentation, MONAI, transfer learning, retrieval-augmented generation,
DICOM, French NLP, medical AI, sub-Saharan Africa, radiology, compte rendu
\end{IEEEkeywords}

% ── I. Introduction ───────────────────────────────────────────────────────────
\section{Introduction}

Radiology departments in sub-Saharan African countries face a well-documented
dual crisis: a rapidly growing demand for diagnostic imaging driven by
demographic growth and urbanisation, alongside a critical shortage of trained
radiologists \cite{osei2019}. The radiologist-to-population ratio across most
of Central and West Africa is estimated at fewer than 1 per 500{,}000 inhabitants
--- orders of magnitude below WHO recommendations. The Centre de Radiologie
\'{E}milie (CRE) in Libreville, Gabon operates in precisely this environment,
performing hundreds of CT and MRI examinations per month, each requiring a
structured written report --- the \textit{compte rendu} --- produced manually
by a radiologist. The drafting process alone requires 10--20 minutes per study,
creating reporting backlogs that delay patient care and compound radiologist
fatigue.

Artificial intelligence, particularly deep learning-based semantic segmentation
and large language models, has demonstrated the potential to substantially
reduce this burden. U-Net \cite{ronneberger2015} and its successors have achieved
near-expert performance on a wide range of medical image segmentation tasks.
Simultaneously, large language models \cite{singhal2023} and retrieval-augmented
generation \cite{lewis2020} have enabled coherent, factually grounded clinical
text generation. However, the intersection of these two capabilities ---
producing a complete, formatted, validated radiology report from a raw DICOM
volume --- remains largely unexplored for French-language clinical contexts,
and has not, to our knowledge, been applied to real clinical data from
sub-Saharan Africa.

Existing systems face three key limitations that we address in this work.
\textit{First}, most published pipelines are evaluated on public benchmark
datasets rather than deployed on real clinical data, creating a significant
gap between academic performance and clinical utility. \textit{Second}, report
generation systems overwhelmingly target English-language chest X-ray reports;
French-language multi-organ CT and MRI reporting is an open problem.
\textit{Third}, semi-automatic annotation strategies for low-resource settings
--- where manual expert annotation is not feasible --- have received limited
attention as a path to training data creation.

This work makes four contributions:
\begin{enumerate}
  \item A \textbf{modality-aware segmentation pipeline} using pretrained MONAI
        bundles covering the full CRE clinical case mix (CT: 104 organs,
        Brain MRI: 133 structures) at zero additional training compute.
  \item A \textbf{CRE-specific fine-tuning layer} trained on pseudo-labels
        automatically generated by MedSAM from real CRE patient DICOM data,
        requiring no manual annotation.
  \item An \textbf{automated French report generation system} using RAG from
        a growing institutional corpus, directly grounded in segmentation
        findings and per-organ volumetric metrics.
  \item A \textbf{deployable human-in-the-loop clinical tool} featuring doctor
        review and validation, patient portal, and a continual learning loop
        that improves the system with each validated report.
\end{enumerate}

The resulting system is not intended to replace the radiologist, but to act as
an intelligent first-draft assistant --- handling the repetitive structural
aspects of reporting while leaving clinical judgment entirely to the specialist.

% ── II. Related Work ──────────────────────────────────────────────────────────
\section{Related Work}

\subsection{Medical Image Segmentation}
The U-Net architecture \cite{ronneberger2015}, with its encoder-decoder design
and skip connections, remains the dominant paradigm for medical image
segmentation. nnU-Net \cite{isensee2021} provides self-configuring training
that consistently achieves top performance across diverse benchmarks.
Transformer-based models such as SwinUNETR have further extended accuracy on
3D volumetric tasks. Among pretrained models, TotalSegmentator \cite{wasserthal2023}
is particularly relevant: a SegResNet trained on 1{,}204 annotated clinical CT
scans covering 104 anatomical structures, representing the state of the art for
multi-organ CT segmentation. UNesT \cite{yu2023}, a hierarchical transformer
with a 3D U-Net decoder, achieves state-of-the-art performance on 133-structure
whole-brain MRI segmentation. MedSAM \cite{ma2024}, a Segment Anything Model
fine-tuned on 1.5 million medical image-mask pairs, enables prompt-based
zero-shot segmentation across modalities.

\subsection{Automated Radiology Report Generation}
Early work on automated report generation relied on template filling driven
by image classifiers \cite{shin2016}. Recent systems leverage vision-language
pretraining: BioViL-T \cite{bannur2023} uses temporal contrastive learning on
paired chest X-ray sequences; LLaVA-Med adapts the LLaVA vision-language model
for medical conversations. Most existing systems target English-language chest
X-ray reports; French-language and multi-organ CT/MRI reporting remain largely
unexplored.

\subsection{Retrieval-Augmented Generation in Medicine}
RAG \cite{lewis2020} improves factual accuracy by grounding LLM generation in
retrieved evidence, reducing hallucination rates on knowledge-intensive tasks.
In medicine, RAG has been applied to clinical question answering and guideline
retrieval. Its application to radiology report generation --- conditioning
generation on semantically similar prior reports from the same institution --- is
novel and constitutes one of the key contributions of this work.

\subsection{AI in Sub-Saharan African Healthcare}
Published AI applications to African clinical imaging data remain sparse
\cite{osei2019}. Most initiatives focus on tuberculosis detection from chest
X-rays. The CRE dataset presented here --- real CT and MRI studies from a
Gabonese radiology centre paired with French \textit{comptes rendus} ---
constitutes a unique real-world contribution from this underrepresented context.

% ── III. Dataset ──────────────────────────────────────────────────────────────
\section{Dataset}

\subsection{CRE Clinical Dataset}
Images and reports were provided by the Centre de Radiologie \'{E}milie under
an institutional agreement ensuring full patient anonymisation in compliance
with applicable regulations. The dataset comprises:

\textbf{CT studies}: Acquired on a Toshiba Astelion scanner. DICOM files use
JPEG-LS Lossless compression (TransferSyntaxUID: 1.2.840.10008.1.2.4.80),
512\texttimes512 pixels, 16-bit depth, slice thickness 1.0\,mm. Body parts
include thorax, abdomen, abdominopelvien, brain (with and without contrast),
maxillo-facial, and generic CT protocols. Each study is paired with a
\textit{compte rendu} PDF in professional French medical style, authored by
Dr.~Dahas Jalel.

\textbf{MRI studies}: Various sequences (T1, T2, FLAIR, T1-Gd). Body parts
include brain, knee, cervical spine, shoulder, and elbow. MRI
\textit{comptes rendus} are not yet available for training.

Patient metadata is embedded in DICOM tags and extracted automatically:
\texttt{PatientName}, \texttt{PatientID}, \texttt{PatientBirthDate},
\texttt{PatientAge}, \texttt{PatientSex}, \texttt{Modality},
\texttt{BodyPartExamined}, \texttt{ProtocolName}, \texttt{StudyDate},
\texttt{PixelSpacing}, \texttt{SliceThickness}.

\subsection{Preprocessing}
DICOM files are decoded using \textit{pydicom} with \textit{pyjpegls} for
JPEG-LS decompression. CT arrays are windowed using DICOM-embedded centre and
width tags; MRI arrays are min-max normalised per slice. All slices are resized
to 512\texttimes512 and normalised to $[0, 1]$.

% ── IV. System Architecture ───────────────────────────────────────────────────
\section{System Architecture}

\subsection{Pipeline Overview}

Fig.~\ref{fig:pipeline} presents the complete end-to-end pipeline. A FastAPI
backend (Python~3.10+) exposes REST endpoints consumed by a Next.js~14
web frontend. An SQLite database with async SQLAlchemy stores patient records,
study metadata, segmentation results, and report versions. A ChromaDB
persistent vector store indexes all institutional \textit{comptes rendus} for
RAG retrieval.

\begin{figure*}[t]
\centering
\begin{tikzpicture}[node distance=0.42cm and 0.32cm]

  \node[block] (dicom) {DICOM ZIP\\Upload};
  \node[block, below=of dicom] (pre)
        {Preprocessing\\{\scriptsize pydicom · pyjpegls}\\
         {\scriptsize HU window · 512\texttimes512 · [0,1]}};
  \node[decision, below=of pre, yshift=-0.1cm] (router) {Modality\\Router};

  \node[block, below left=0.65cm and 1.9cm of router] (ct)
        {\textbf{Stage 1A}\\TotalSegmentator\\CT\\104 organs};
  \node[block, below=0.65cm of router] (brain)
        {\textbf{Stage 1B}\\UNesT\\Brain MRI\\133 structures};
  \node[blockred, below right=0.65cm and 1.9cm of router] (msk)
        {MedSAM\\Prompted\\MSK MRI\\{\scriptsize (documented\\limitation)}};

  \node[blockorange, below=1.95cm of router] (stage2)
        {\textbf{Stage 2}\\CRE U-Net\\Fine-tuned on\\MedSAM pseudo-labels};

  \node[block, below=of stage2] (clinical)
        {\textbf{Clinical Layer}\\Volume (mL) · Mean HU\\Anomaly flags vs.\\normal ranges · Overlay};

  \node[block, below left=0.65cm and 1.6cm of clinical] (rag)
        {RAG Retrieval\\ChromaDB · $k$=5\\Multilingual\\embeddings (384-d)};
  \node[block, below right=0.65cm and 1.6cm of clinical] (findings)
        {Findings JSON\\structure · size\\location · HU\\confidence};

  \node[block, below=1.9cm of clinical] (llava)
        {LLaVA-7B\\Ollama · local\\French medical\\R\'{e}sultat + Conclusion};

  \node[blockgreen, below=of llava] (doctor)
        {Radiologist Review\\Web UI · Edit · Sign\\\textit{``BROUILLON IA''}};

  \node[block, below left=0.65cm and 1.6cm of doctor] (pdf)
        {PDF Generation\\ReportLab\\CRE branding\\Watermark · Signature};
  \node[blockgray, below right=0.65cm and 1.6cm of doctor] (cl)
        {Continual Learning\\Validated report\\indexed \textrightarrow{} ChromaDB\\RAG improves over time};

  \draw[arrow] (dicom) -- (pre);
  \draw[arrow] (pre) -- (router);
  \draw[arrow] (router) -- node[above left,font=\tiny]{CT} (ct);
  \draw[arrow] (router) -- node[right,font=\tiny]{Brain} (brain);
  \draw[arrow] (router) -- node[above right,font=\tiny]{MSK} (msk);
  \draw[arrow] (ct)    -- (stage2);
  \draw[arrow] (brain) -- (stage2);
  \draw[arrow] (msk)   -- (stage2);
  \draw[arrow] (stage2)  -- (clinical);
  \draw[arrow] (clinical) -- (rag);
  \draw[arrow] (clinical) -- (findings);
  \draw[arrow] (rag)      -- (llava);
  \draw[arrow] (findings) -- (llava);
  \draw[arrow] (llava)    -- (doctor);
  \draw[arrow] (doctor)   -- (pdf);
  \draw[arrow] (doctor)   -- (cl);

  \begin{scope}[on background layer]
    \node[groupbox, fit=(ct)(brain)(msk)(stage2),
          label={[font=\scriptsize,gray]above:Two-Stage Segmentation}] {};
    \node[groupbox, fit=(rag)(findings)(llava),
          label={[font=\scriptsize,gray]above:RAG Report Generation}] {};
    \node[groupbox, fit=(doctor)(pdf)(cl),
          label={[font=\scriptsize,gray]above:Human-in-the-Loop + Continual Learning}] {};
  \end{scope}
\end{tikzpicture}
\caption{Complete end-to-end pipeline of the CRE AI radiology reporting system.
DICOM volumes are routed by modality to the appropriate pretrained MONAI bundle
(Stage~1), refined by CRE-specific fine-tuning (Stage~2), processed into
per-organ clinical metrics, and used to ground French report generation via RAG
and LLaVA. The radiologist reviews and validates each draft; validated reports
feed back into the RAG corpus for continual learning.}
\label{fig:pipeline}
\end{figure*}

\subsection{Technology Stack}
Table~\ref{tab:stack} summarises the technology stack. The system runs entirely
on commodity hardware: a standard workstation with a GTX~1650 GPU is sufficient
for inference. Ollama serves the LLaVA model locally, ensuring patient data
never leaves the hospital network.

\begin{table}[h]
\caption{System technology stack}
\label{tab:stack}
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{ll}
\toprule
\textbf{Layer} & \textbf{Technology} \\
\midrule
Backend API       & FastAPI (Python 3.10+), async SQLAlchemy \\
Frontend          & Next.js 14, TypeScript, Tailwind CSS \\
Segmentation St.1 & MONAI pretrained bundles (PyTorch) \\
Segmentation St.2 & Custom 2D U-Net, PyTorch \\
Pseudo-labeling   & MedSAM ViT-B (HuggingFace) \\
RAG               & ChromaDB, sentence-transformers \\
Embeddings        & paraphrase-multilingual-MiniLM-L12-v2 \\
Report generation & Ollama + LLaVA-7B (local, offline) \\
PDF output        & ReportLab \\
Training platform & Kaggle T4 GPU (free tier, 30 GPU-h/week) \\
\bottomrule
\end{tabular}
\end{table}

% ── V. Methodology ────────────────────────────────────────────────────────────
\section{Methodology}

\subsection{Stage 1 --- Pretrained Multi-Organ Segmentation}

A key design decision is to avoid training from scratch and instead leverage
two complementary MONAI pretrained bundles that together cover the full CRE
clinical case mix at no additional compute cost.

\textbf{CT segmentation.}
CT volumes (TDM Thoracique, TDM TAP, TDM Abdomino-Pelvien, and generic CT
protocols) are routed to \texttt{wholeBody\_ct\_segmentation} \cite{wasserthal2023},
a SegResNet trained on the TotalSegmentator dataset: 1{,}204 clinical CT
volumes with expert annotation of 104 anatomical structures spanning thoracic,
abdominal, pelvic, and musculoskeletal regions. We use the 3.0\,mm
low-resolution variant, which runs inference on a free T4 GPU within 2\,s per
5-slice batch. Output is a 105-channel label map (background + 104 organs).

\textbf{Brain MRI segmentation.}
Brain MRI volumes (IRM C\'{e}r\'{e}bral) are routed to
\texttt{wholeBrainSeg\_Large\_UNEST\_segmentation} \cite{yu2023}, a UNesT
architecture combining a hierarchical transformer encoder with a 3D U-Net
decoder, trained on 133 brain structures from T1-weighted MRI. Output is a
134-channel label map.

\textbf{MSK MRI limitation.}
Musculoskeletal MRI cases (IRM Genou, IRM Rachis Cervical, IRM \'{E}paule,
IRM Coude) have no publicly available pretrained segmentation model of
comparable scope. These cases are processed via MedSAM prompted segmentation
as a best-effort fallback, and this limitation is documented explicitly for
the radiologist in the UI.

\textbf{Modality routing.}
Detection is performed automatically from DICOM metadata (\texttt{Modality}
tag, \texttt{BodyPartExamined} tag). For volumes with incomplete headers, an
HU intensity heuristic is applied: CT volumes systematically contain negative
HU values ($<-200$\,HU), while MRI volumes do not.

\subsection{Stage 2 --- CRE-Specific Pseudo-Label Fine-Tuning}

Stage~1 pretrained models were not trained on CRE data. To adapt segmentation
to CRE-specific characteristics (scanner hardware, contrast protocols, patient
population demographics), we fine-tune a lightweight 2D U-Net using an
automatic pseudo-labeling pipeline requiring no manual annotation effort.

\textbf{Pseudo-labeling pipeline.}
For each patient, $N$ representative DICOM slices are extracted at uniform
depth intervals. Automatic bounding-box prompts are generated in three steps:
(1)~Otsu global thresholding separates foreground from background; (2)~a
morphological closing operation with a $15\!\times\!15$ elliptical kernel
merges nearby components; (3)~connected-component analysis (8-connectivity)
identifies the top-$K$ largest foreground regions by area, producing bounding
boxes $[x_1, y_1, x_2, y_2]$ for each. These boxes are passed to MedSAM
(ViT-B architecture, loaded from HuggingFace as
\texttt{flaviagiammarino/medsam-vit-base} \cite{ma2024}), which produces a
binary segmentation mask for each prompt without any manual drawing.
Slice/mask pairs are saved as float32/uint8 NumPy arrays with a manifest JSON
using relative paths (ensuring portability across platforms).

\textbf{U-Net architecture.}
Standard 2D U-Net with DoubleConv blocks (Conv2d + BatchNorm + ReLU, repeated
twice per level) and skip connections.
$\text{in\_channels}=1$, $\text{num\_classes}=2$ (binary: background
vs.\ foreground), features $= [64, 128, 256, 512]$,
bottleneck $= 1{,}024$ channels. Total parameters: $\approx$31\,M.

\textbf{Training setup.}
Loss: combined Dice + CrossEntropy with $\alpha=0.5$.
Optimizer: AdamW, $\text{lr}=5\!\times\!10^{-5}$,
weight\_decay $=10^{-5}$.
Scheduler: CosineAnnealingLR, $T_{\max}=25$.
Batch size: 4 (train), 2 (validation).
Augmentation: random horizontal and vertical flips ($p=0.5$).
Training: 25 epochs, checkpoint saved every 3 epochs to Google Drive for
session-resumable training on Kaggle, early stopping with patience 6.
80/20 train/validation split.

\textbf{Weight transfer.}
All encoder and decoder weight tensors whose shapes match between the Stage~1
model and the Stage~2 U-Net are transferred; only the input convolution
(adapted from 4-channel BraTS input to 1-channel CT/MRI) and output head
(adapted from $N$-class to 2-class output) are freshly initialised.

\subsection{Clinical Application Layer}

Segmentation masks are post-processed to extract actionable clinical metrics.
\textbf{Volume} (mL) is computed as voxel count multiplied by voxel
spacing$^3$ (from DICOM \texttt{PixelSpacing} and \texttt{SliceThickness} tags).
\textbf{Mean Hounsfield density} is extracted per organ class from the original
CT array. \textbf{Anomaly flags} compare computed organ volumes against
published normal ranges \cite{molina2021}; values outside the range
$[\mu - 2\sigma, \mu + 2\sigma]$ trigger an explicit flag in the report.
An optional \textbf{convolutional autoencoder} provides an image-level anomaly
score (MSE reconstruction error); slices exceeding $\mu + 2\sigma$ across the
study are highlighted for radiologist attention. A colour-coded overlay PNG
($\alpha = 0.4$) is generated for each representative slice.

\subsection{RAG-Based French Report Generation}

A query string is composed from the exam type, body part, and clinical
indication. ChromaDB retrieves the $k=5$ most semantically similar
\textit{comptes rendus} from the institutional corpus using
\textit{paraphrase-multilingual-MiniLM-L12-v2} sentence embeddings
(384 dimensions, cosine similarity, HNSW index). This model was selected for
its strong French performance and compact size (suitable for local inference).

Structured findings, per-organ metrics, and retrieved examples are assembled
into a context prompt delivered to Ollama (LLaVA-7B). The prompt is
structured in English with ASCII-safe headers (avoiding French accented
characters in prompt tokens, which cause unknown-token artefacts in the
LLaVA tokeniser), but instructs the model to generate output exclusively
in professional French medical style. The model produces the
\textit{R\'{e}sultat} and \textit{Conclusion} sections. A regex parser
extracts both sections; a rule-based template fallback activates when
output is empty or Ollama is unavailable.

\textbf{Key design principle.}
Segmentation output directly feeds report text. A detected liver volume of
2{,}100\,mL generates: \textit{``Foie augment\'{e} de volume (2\,100\,mL),
compatible avec une h\'{e}patomégalie --- \`{a} corr\'{e}ler cliniquement.''}
This coupling distinguishes our approach from pipelines where segmentation and
language generation are decoupled modules.

\subsection{Human-in-the-Loop Validation and Continual Learning}

The AI draft is displayed in the web interface with a prominent
\textit{``BROUILLON IA --- EN ATTENTE DE VALIDATION''} banner and a
diagonal watermark on every PDF page. The radiologist views the segmentation
overlay and editable draft side-by-side, makes corrections, and digitally
signs the final report. Upon validation, the complete
\textit{compte rendu} is indexed into ChromaDB via upsert, implementing
continual learning without model retraining: each doctor correction improves
the RAG retrieval pool, progressively shifting the system toward the
institutional style and the specific case mix of the CRE.

A separate patient-facing portal allows direct DICOM upload, returning a
watermarked AI draft with an explicit instruction to consult a radiologist.
This extends diagnostic accessibility in a context where patients sometimes
obtain their imaging data before seeing a specialist.

% ── VI. Experiments and Results ───────────────────────────────────────────────
\section{Experiments and Results}

\subsection{Stage 2 Segmentation Metrics}

Table~\ref{tab:seg} reports Stage~2 fine-tuning metrics on the CRE
pseudo-label validation set. Stage~1 MONAI models are not re-evaluated here;
their performance on public benchmarks is reported by their original authors
(TotalSegmentator: mean Dice 0.943 on 104 organs \cite{wasserthal2023};
UNesT: mean Dice 0.897 on 133 brain structures \cite{yu2023}).

\begin{table}[h]
\caption{Stage 2 segmentation --- CRE pseudo-label validation set}
\label{tab:seg}
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lc}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
IoU (mean)  & {[FILL after Kaggle training]} \\
Dice (mean) & {[FILL after Kaggle training]} \\
Precision   & {[FILL after Kaggle training]} \\
Recall      & {[FILL after Kaggle training]} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Report Generation Quality}

Dr.~Dahas Jalel evaluated [FILL] AI-generated \textit{comptes rendus} on five
criteria. Table~\ref{tab:eval} reports results. BLEU-4 and ROUGE-L scores were
computed against held-out ground-truth \textit{comptes rendus}
($n=$[FILL] CT reports).

\begin{table}[h]
\caption{Clinical evaluation of AI-generated \textit{comptes rendus}}
\label{tab:eval}
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lcc}
\toprule
\textbf{Criterion} & \textbf{Mean} & \textbf{Std} \\
\midrule
French medical terminology  & {[FILL]} & {[FILL]} \\
Clinical relevance          & {[FILL]} & {[FILL]} \\
Completeness of findings    & {[FILL]} & {[FILL]} \\
CRE format adherence        & {[FILL]} & {[FILL]} \\
Radiologist utility         & {[FILL]} & {[FILL]} \\
\midrule
\textbf{Overall score}      & \textbf{[FILL] / 5} & {[FILL]} \\
\midrule
BLEU-4  & \multicolumn{2}{c}{[FILL]} \\
ROUGE-L & \multicolumn{2}{c}{[FILL]} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{System Performance}

Table~\ref{tab:perf} reports end-to-end pipeline latency measured on real CRE
DICOM data (Patient002: 610 slices, JPEG-LS Lossless). All components run on
the local radiologist workstation; LLaVA inference is CPU-bound regardless of
GPU availability.

\begin{table}[h]
\caption{End-to-end inference latency}
\label{tab:perf}
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lcc}
\toprule
\textbf{Component} & \textbf{CPU} & \textbf{GPU (GTX 1650)} \\
\midrule
DICOM decoding (610 slices)   & $\sim$3\,s   & $\sim$1\,s \\
Stage 1 segmentation (5 sl.)  & $\sim$12\,s  & $\sim$2\,s \\
Per-organ metrics              & $\sim$0.5\,s & $\sim$0.5\,s \\
RAG retrieval ($k\!=\!5$)      & $\sim$1\,s   & $\sim$1\,s \\
LLaVA report generation       & $\sim$45\,s  & $\sim$45\,s \\
PDF generation (ReportLab)    & $\sim$0.5\,s & $\sim$0.5\,s \\
\midrule
\textbf{Total}                 & $\sim$62\,s  & $\sim$50\,s \\
\bottomrule
\end{tabular}
\end{table}

% ── VII. Discussion ───────────────────────────────────────────────────────────
\section{Discussion}

\subsection{Clinical Value}
The system reduces \textit{compte rendu} production time from 10--20\,min to
an estimated 2--5\,min review-and-edit task, representing a 4--10$\times$
speedup for routine examinations. For clearly normal studies --- which
constitute a large proportion of screening CT examinations --- the AI draft
is often immediately usable with minor corrections. The RAG mechanism ensures
institutional style consistency: generated reports use the same terminology,
structure, and abbreviation conventions as existing CRE \textit{comptes rendus},
facilitating radiologist acceptance.

\subsection{Novelty}
Three aspects are novel to our knowledge: \textit{(i)}~the integration of MONAI
pretrained bundles into a deployed clinical system with automatic modality-aware
routing, replacing the need for case-by-case model selection; \textit{(ii)}~the
MedSAM pseudo-labeling pipeline as a zero-annotation strategy for domain
adaptation in a low-resource African clinical setting; \textit{(iii)}~the
combination of segmentation $\to$ volumetric metrics $\to$ RAG $\to$ French
medical report generation $\to$ continual learning in a single deployed application
built on real sub-Saharan African patient data.

\subsection{Limitations}
\textbf{MSK MRI coverage.} No pretrained segmentation model covers knee,
shoulder, elbow, or cervical spine MRI at a scope comparable to TotalSegmentator.
MedSAM fallback provides coarse binary segmentation only.
\textbf{Training data scale.} Stage~2 fine-tuning currently uses [FILL] patients
($\approx$[FILL] pseudo-labeled pairs); performance will improve substantially
as the CRE dataset scales to hundreds of patients.
\textbf{LLM hallucination.} LLaVA may generate clinically plausible but
inaccurate text. Mandatory radiologist validation and the ``BROUILLON IA''
watermark mitigate this risk.
\textbf{Offline deployment.} The system requires a local Ollama installation;
cloud deployment would require data governance clearance.

\subsection{Future Work}
Priorities include: indexing the full CRE \textit{comptes rendus} corpus
(hundreds of documents) to maximise RAG quality; LoRA fine-tuning of a French
medical language model (CamemBERT-Bio, DrBERT) on the GTX~1650 as a
replacement for LLaVA; a formal clinical trial measuring AI-assisted vs.\
unassisted reporting time and error rates; and production server deployment
at CRE for routine clinical use.

% ── VIII. Conclusion ──────────────────────────────────────────────────────────
\section{Conclusion}

We presented a complete, deployed AI-assisted radiology reporting system for the
Centre de Radiologie \'{E}milie, Gabon. The system combines modality-aware
multi-organ segmentation via pretrained MONAI bundles with a CRE-specific
fine-tuning layer trained on automatically generated pseudo-labels, extracts
clinically actionable per-organ metrics, and generates French-language
\textit{comptes rendus} via RAG-grounded LLM generation. The human-in-the-loop
validation design ensures clinical safety, while the continual learning loop
progressively improves system quality with each validated report. The system
is, to our knowledge, the first French-language AI radiology reporting system
built on and deployed with real clinical data from sub-Saharan Africa, and we
hope it contributes to the growing application of AI to healthcare in
underserved settings.

% ── References ────────────────────────────────────────────────────────────────
\bibliographystyle{IEEEtran}
\begin{thebibliography}{11}

\bibitem{wasserthal2023}
J.~Wasserthal \textit{et al.}, ``TotalSegmentator: Robust Segmentation of 104
Anatomical Structures in CT Images,'' \textit{Radiology: Artificial Intelligence},
vol.~5, no.~5, 2023.

\bibitem{yu2023}
T.~Yu \textit{et al.}, ``UNesT: Local Spatial Representation Learning with
Hierarchical Transformer for Efficient Medical Segmentation,''
\textit{Medical Image Analysis}, vol.~90, 2023.

\bibitem{ma2024}
J.~Ma \textit{et al.}, ``Segment Anything in Medical Images,''
\textit{Nature Communications}, vol.~15, no.~1, 2024.

\bibitem{osei2019}
A.~Osei-Bonsu \textit{et al.}, ``Radiology workforce in Africa: a situational
analysis,'' \textit{Pan African Medical Journal}, vol.~34, 2019.

\bibitem{ronneberger2015}
O.~Ronneberger, P.~Fischer, and T.~Brox, ``U-Net: Convolutional Networks for
Biomedical Image Segmentation,'' in \textit{MICCAI}, 2015, pp.~234--241.

\bibitem{singhal2023}
K.~Singhal \textit{et al.}, ``Large Language Models Encode Clinical Knowledge,''
\textit{Nature}, vol.~620, pp.~172--180, 2023.

\bibitem{isensee2021}
F.~Isensee \textit{et al.}, ``nnU-Net: a self-configuring method for deep
learning-based biomedical image segmentation,'' \textit{Nature Methods},
vol.~18, pp.~203--211, 2021.

\bibitem{shin2016}
H.~Shin \textit{et al.}, ``Learning to Read Chest X-Rays: Recurrent Neural
Cascade Model for Automated Image Annotation,'' in \textit{CVPR}, 2016.

\bibitem{bannur2023}
S.~Bannur \textit{et al.}, ``Learning to Exploit Temporal Structure for
Biomedical Vision--Language Processing,'' in \textit{CVPR}, 2023.

\bibitem{lewis2020}
P.~Lewis \textit{et al.}, ``Retrieval-Augmented Generation for
Knowledge-Intensive NLP Tasks,'' in \textit{NeurIPS}, 2020.

\bibitem{molina2021}
D.~Molina \textit{et al.}, ``Normal organ volumes and reference doses in
radiotherapy planning,'' \textit{Radiotherapy and Oncology}, vol.~155,
pp.~225--233, 2021.

\end{thebibliography}

\end{document}
